---
output: 
  github_document:
    fig_width: 6
    fig_heigh: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, eval=TRUE,  cache=TRUE)
library(Rata)
```

# Rata: Automated Test Assembly 


## Overview

Rata applies mixed integer programming (MIP) to automatically build test forms for educational and psychological assessment. Rata is designed as a general automated test assembly (ATA) framework, and it can be used to assemble linear tests, support linear-on-the-fly testing (LOFT), prepare item pools for computerized adaptive testing (CAT), and assemble panels in multistage testing (MST). 

In Rata, users first define a MIP model by adding the objective and constraints sequentially, and then solve the MIP model using either [lp_solve](http://lpsolve.sourceforge.net/5.5/) or [GLPK](https://www.gnu.org/software/glpk/).

Rata supports mixed format item pool. Currently, the mixed format item pool supports a mixture of 3PL, GPCM, and GRM items. Key dependencies of Rata include [Rirt](https://CRAN.R-project.org/package=Rirt), [lpSolveAPI](https://CRAN.R-project.org/package=lpSolveAPI), and [glpkAPI](https://CRAN.R-project.org/package=glpkAPI). 

## Installation

Install the stable version from CRAN:
```{r, eval=FALSE}
install.packages("Rata")
```

Install the most recent version from [github](https://github.com/xluo11/Rata):
```{r, eval=FALSE}
devtools::install_github("xluo11/Rata")
```


## Quickstart

Users would use the following functions to build and solve MIP models:

+ `ata`: initiate a MIP model
+ `ata_relative_objective`: add relative test assembly objective to the model
+ `ata_absolute_objective`: add absolute test assembly objective to the model
+ `ata_constraint`: add constraints pertaining to categorical and continuous item attributes to the model
+ `ata_item_use`: add constraints pertaining to item usage to the model
+ `ata_item_enemy`: add constraints pertaining to enemy items to the model
+ `ata_item_fix`: fix the value of the decision variable in the model
+ `ata_solve`: solve the model using a MIP solver. Rata currently supports lp_solve and GLPK

## Usage

Write a helper function for generating item pools which will be used in following examples. By default, the pool includes 200 3PL items, 20 GPCM items, and 20 GRM items. Each item has a categorical item attribute (content area) and a continuous attribute (response time). In addition, items are randomly grouped to form item sets.

```{r}
library(Rirt)
library(dplyr, warn.conflicts=FALSE)

item_pool <- function(types=c('3pl', 'gpcm', 'grm'), n_3pl=300, n_gpcm=20, n_grm=20, seed=21578) {
  set.seed(seed)
  items <- model_mixed_gendata(1, n_3pl, n_gpcm, n_grm, n_c=4)$items
  items$'3pl' <- cbind(items$'3pl', id=seq(n_3pl), content=sample(3, n_3pl, replace=T),
                       time=round(rlnorm(n_3pl, 4, .38)), group=sort(sample(n_3pl/2, n_3pl, replace=T)))
  items$'gpcm' <- cbind(items$'gpcm', id=seq(n_gpcm), content=sample(3, n_gpcm, replace=T),
                        time=round(rlnorm(n_gpcm, 4, .38)), group=sort(sample(n_gpcm/2, n_gpcm, replace=T)))
  items$'grm' <- cbind(items$'grm', id=seq(n_grm), content=sample(3, n_grm, replace=T),
                       time=round(rlnorm(n_grm, 4, .38)), group=sort(sample(n_grm/2, n_grm, replace=T)))
  items[names(items) %in% types]
}
```

#### Example 1: Assemble linear tests

Assemble 4 non-overlapping forms that should have maximum test information function (TIF) in the region [-1.28, 1.28] on the scale which covers 80% of the population. Each form includes 10 items, with 3, 3, and 4 items in the content area 1 to 3 respectively. Additionally, each form has an average response time of 60 +/- 5 seconds per item.

```{r}
x <- ata(item_pool(), n_forms=4, test_len=10, max_use=1)
x <- ata_relative_objective(x, seq(-1.28, 1.28, length.out=4), 'max')
x <- ata_constraint(x, 'content', min=3, max=3, level=1)
x <- ata_constraint(x, 'content', min=3, max=3, level=2)
x <- ata_constraint(x, 'content', min=4, max=4, level=3)
x <- ata_constraint(x, 'time', min=55*10, max=65*10)
x <- ata_solve(x, 'lpsolve', return_format='model', time_limit=30)
rbind(if(nrow(x$items$'3pl') >  0) select(x$items$'3pl', content, time, form) else NULL,
      if(nrow(x$items$'gpcm') > 0) select(x$items$'gpcm', content, time, form) else NULL,
      if(nrow(x$items$'grm') >  0) select(x$items$'grm', content, time, form) else NULL) %>%
  group_by(form) %>%
  summarise(con1=sum(content==1), con2=sum(content==2), con3=sum(content==3), time=mean(time))
plot(x)
```

To obtain a flatter TIF over the central region of the scale that produces more comparable measurement errrors, we give ATA a TIF target that is about 60% of the optimum found in the previous solution.

```{r}
tif_tar <- x$obj_var[1] * .6
x <- ata(item_pool(), n_forms=4, test_len=10, max_use=1)
x <- ata_absolute_objective(x, seq(-1.28, 1.28, length.out=4), target=tif_tar)
x <- ata_constraint(x, 'content', min=3, max=3, level=1)
x <- ata_constraint(x, 'content', min=3, max=3, level=2)
x <- ata_constraint(x, 'content', min=4, max=4, level=3)
x <- ata_constraint(x, 'time', min=55*10, max=65*10)
x <- ata_solve(x, 'lpsolve', return_format='model', time_limit=30)
rbind(if(nrow(x$items$'3pl') >  0) select(x$items$'3pl', content, time, form) else NULL,
      if(nrow(x$items$'gpcm') > 0) select(x$items$'gpcm', content, time, form) else NULL,
      if(nrow(x$items$'grm') >  0) select(x$items$'grm', content, time, form) else NULL) %>%
  group_by(form) %>%
  summarise(con1=sum(content==1), con2=sum(content==2), con3=sum(content==3), time=mean(time))
plot(x)
```

We can also use difficulty parameters, as opposed to TIF, to control the psychometric characteristics of the test forms. In this example, we assemble four parallel test forms in which the mean and SD of difficulty parameters are equal to 0 and 1 respectively.

```{r}
x <- ata(item_pool('3pl', n_3pl=300), n_forms=4, test_len=10, max_use=1) # generate 300 3PL items
x <- ata_absolute_objective(x, x$pool$'3pl'$b, target=0*10)
x <- ata_absolute_objective(x, (x$pool$'3pl'$b-0)^2, target=1^2*10)
x <- ata_constraint(x, 'content', min=3, max=3, level=1)
x <- ata_constraint(x, 'content', min=3, max=3, level=2)
x <- ata_constraint(x, 'content', min=4, max=4, level=3)
x <- ata_constraint(x, 'time', min=55*10, max=65*10)
x <- ata_solve(x, 'lpsolve', return_format='model', time_limit=30)
group_by(x$items$'3pl', form) %>%
  summarise(b_mean=mean(b), b_std=sd(b), con1=sum(content==1), con2=sum(content==2), con3=sum(content==3), time=mean(time)) %>% round(., 2)
```


#### Example 2: Model item relational constraints

In this example, we demonstrate how to control item relational constraints in ATA. First is the item set constraints, which select items associated with the same stimulus as a group. Simply give the item set grouping variable to the `group` argument in the `ata` function would add that constraint to the MIP model.

```{r}
x <- ata(item_pool('3pl', n_3pl=200), n_forms=4, test_len=10, max_use=1, group='group') # generate 200 3PL items
x <- ata_relative_objective(x, seq(-.68, .68, length.out=3), 'max') # maximize TIF over [-.68, .68] which covers 50% of the population
x <- ata_solve(x, 'lpsolve', 'model', time_limit=30)
# ATA = number of items associated with the group id in assembled forms
# Pool = number of items associated with the group id in the item pool
merge(group_by(x$items$'3pl', form, group) %>% summarise(ATA=n()),
       group_by(x$pool$'3pl', group) %>% summarise(Pool=n()),
       by='group', all.x=TRUE) %>% arrange(form, group)
plot(x)
```

To have a set of common items shared in all forms, use `ata(..., common_items=)`. To have a set of common items shared between adjacent forms, use `ata(..., overlap_items)`.

```{r}
x <- ata(item_pool('3pl', n_3pl=200), n_forms=4, test_len=10, max_use=1, common_items=5)
x <- ata_relative_objective(x, seq(-.68, .68, length.out=3), 'max')
x <- ata_solve(x, 'lpsolve', 'model', time_limit=30)
plot(x)
counts <- with(x$items$'3pl', table(id, form))
# expect 10 items each test form
colSums(counts)
# expect 5 items that appear in all test forms
counts[rowSums(counts)==4, ]
```

To avoid enemy items in the same pool, use the `ata_item_enemy` function.

```{r}
x <- ata(item_pool('3pl', n_3pl=15), n_forms=5, test_len=3, max_use=1)
x <- ata_relative_objective(x, x$pool$'3pl'$b, 'max')
x <- ata_item_enemy(x, 1:5) # do not put item 1-5 in the same form
x <- ata_item_enemy(x, 6:10) # do not put item 6-10 in the same form
x <- ata_item_enemy(x, 11:15) # do not put item 11-15 in the same form
x <- ata_solve(x, 'lpsolve', 'model', time_limit=30)
# expect each form have one item from 1-5, 6-10, and 11-15
select(x$items$'3pl', form, id) %>% arrange(form, id)
```

To allow item reuse, use `ata(..., max_use=)` or the `ata_item_use` function.

```{r}
x <- ata(item_pool('3pl', n_3pl=15), n_forms=2, test_len=5, max_use=2) # allow items to be used up to twice
x <- ata_relative_objective(x, x$pool$'3pl'$b, 'max')
x <- ata_solve(x, 'lpsolve', 'model', time_limit=30)
# expect two identical forms, since the best items can be reused
with(x$items$'3pl', table(id, form))
```

To force an item to be select or not selected, use the `` function.

```{r}
x <- ata(item_pool('3pl', n_3pl=15), n_forms=2, test_len=5, max_use=2)
x <- ata_relative_objective(x, x$pool$'3pl'$b, 'max')
x <- ata_item_fix(x, 1:5, min=1, forms=1) # use item 1-5 in form 1
x <- ata_item_fix(x, 1:5, max=0, forms=2) # do not use 1-5 in form 2
x <- ata_solve(x, 'lpsolve', 'model', time_limit=30)
# expect item 1-5 in form 1 but not form 2
with(x$items$'3pl', table(id, form))
```


## Getting help

If you encounter a bug, please post a code example that exposes the bug on [github](https://github.com/xluo11/Rata/issues). You can post your questions and feature requests on [github](https://github.com/xluo11/Rata/issues) or to the [author](mailto:xluo1986@gmail.com).
